---
title: "Responding to Reviewer Comments"
author: "Will Townes"
date: "8/19/2020"
output: html_document
---

```{r}
library(tidyverse); theme_set(theme_bw())
```

## Reviewer 1

### Comment 3

Could the authors provide a visual representation of the overlap (or lack thereof) between predicted pro- and anti-longevity genes predicted from GO only vs. GO+ARCHS2 (mentioned to differ quite a lot in manuscript p4). This would be best done by maybe comparing ranks as a scatterplot, or a Venn diagram with more than 10 genes on each side. This would give an idea of how both types of models perform and predict.

```{r}
#GO+ARCHS4 features
worm<-read.table("./results/final_preds_go_archs4_worm.txt",header=TRUE)
worm$species<-"worm"
yeast<-read.table("./results/final_preds_go_archs4_yeast.txt",header=TRUE)
yeast$species<-"yeast"
pd<-rbind(worm,yeast)
colnames(pd)[colnames(pd)=="prob_pro_longevity"]<-"go_archs4_prob"
#GO only features
worm<-read.table("./results/final_preds_go_only_worm.txt",header=TRUE)
worm$species<-"worm"
yeast<-read.table("./results/final_preds_go_only_yeast.txt",header=TRUE)
yeast$species<-"yeast"
pd2<-rbind(worm,yeast)
colnames(pd2)[colnames(pd2)=="prob_pro_longevity"]<-"go_only_prob"
pd<-merge(pd,pd2,by=c("species","gene","genage_annotated"))
pd$ga<-"non-genage"
pd$ga[pd$genage_annotated]<-"genage"
ggplot(pd,aes(x=go_archs4_prob,y=go_only_prob))+geom_point()+facet_grid(ga~species)
ggsave("./results/plots/probs_compare_go_archs4_vs_go_only.pdf",width=5,height=5)

#same plot but using ranks- too noisy
scaled_rank<-function(x){rank(x,na.last="keep")/length(x)}
pd2<-pd %>% group_by(species) %>%
  mutate(go_archs4_rk=scaled_rank(go_archs4_prob),
         go_only_rk=scaled_rank(go_only_prob))
ggplot(pd2,aes(x=go_archs4_rk,y=go_only_rk))+geom_point()+facet_grid(ga~species)
ggsave("./results/plots/probs_compare_go_archs4_vs_go_only_ranks.pdf",width=5,height=5)
```